---
title: "Assignment 4"
author: "Venkata Naga Siddartha Gutha"
output: pdf_document
---
Loading libraries and data set
```{r}
library(tidyverse)
library(factoextra)
pharmaceutical_data<-read.csv("C:/Users/sidda/Downloads/Pharmaceuticals.csv")
pharmaceutical_data<-na.omit(pharmaceutical_data)

```

Using the numerical variables (1 to 9) to cluster the 21 firms.
```{r}
row.names(pharmaceutical_data)<-pharmaceutical_data[,1]
Clustering_dataset<-pharmaceutical_data[,3:11]
```
Scaling the data
```{r}
set.seed(143)
Scaled_data<-scale(Clustering_dataset)

```

Performing Kmeans for random K values
```{r}
set.seed(143)
kmeans_2<-kmeans(Scaled_data,centers = 2, nstart = 15)
kmeans_4<-kmeans(Scaled_data,centers = 4, nstart = 15)
kmeans_8<-kmeans(Scaled_data,centers = 8, nstart = 15)

plot_kmeans_2<-fviz_cluster(kmeans_2,data = Scaled_data) + ggtitle("K=2")
plot_kmeans_4<-fviz_cluster(kmeans_4,data = Scaled_data) + ggtitle("K=4")
plot_kmeans_8<-fviz_cluster(kmeans_8,data = Scaled_data) + ggtitle("K=8")

plot_kmeans_2
plot_kmeans_4
plot_kmeans_8


```

Using WSS and Silhouette to find best K suitable for clustering

```{r}
k_wss<-fviz_nbclust(Scaled_data,kmeans,method="wss")
k_silhouette<-fviz_nbclust(Scaled_data,kmeans,method="silhouette")
k_wss
k_silhouette
distance<-dist(Scaled_data,metho='euclidean')
fviz_dist(distance)
```
from WSS k is 2 and from silhouette k is 5. we are choosing 5 as this ensures that within sum of squires is low along with good separation within clusters

Performing Kmeans for suitable k 
```{r}
set.seed(143)
kmeans_5<-kmeans(Scaled_data,centers = 5, nstart = 10)
kmeans_5
plot_kmeans_5<-fviz_cluster(kmeans_5,data = Scaled_data) + ggtitle("K=5")
plot_kmeans_5

Clustering_dataset_1<-Clustering_dataset%>%
  mutate(Cluster_no=kmeans_5$cluster)%>%
  group_by(Cluster_no)%>%summarise_all('mean')
Clustering_dataset_1

```
Companies are grouped into following clusters:


Cluster_1= ABT,AHM,AZN,BMY,LLY,NVS,SGP,WYE


Cluster_2= BAY,CHTT,IVX


Cluster_3=AVE,ELN,MRX,WPI


Cluster_4=AGN,PHA


Cluster_5=GSK,JNJ,MRK,PFE




From the clusters formed it can be understood that 


1. Cluster_1 has group of companies with moderate return on equity and return on investment


2. Cluster_2 contains companies with very bad ROA,ROE, market capitalization and asset turnover. this implies that these companies are very risky


3. Cluster_3 has group companies similar to cluster_2 but with little less risk involved


4. Cluster_4 companies has very good PE_ratio but very poor ROA,ROE which is more riskier that cluster_2


5. Cluster_5 has companies with very good market capitalization, ROE and ROA


```{r}
Clustering_datase_2<- pharmaceutical_data[,12:14] %>% mutate(Clusters=kmeans_5$cluster)
ggplot(Clustering_datase_2, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')
ggplot(Clustering_datase_2, mapping = aes(factor(Clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='Clusters')
ggplot(Clustering_datase_2, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='Clusters')


```

It can be seen that there is a pattern in clusters  and the variable Median Recommendation. Like the 2nd cluster suggests between hold and moderate buy,3rd cluster suggests to moderate buy to moderate sell.
From the location graph it can be noticed that most of the pharmaceutical companies are US based and there is no much pattern in it.
There is no noticeable pattern between clusters and exchange except the fact that majority of companies are listed on NYSE.


Naming clusters:


[It is done based net Market capitalization(size) and Return on Assets(money)]


Cluster 1: Large-Thousands


Cluster 2: Extra Small-Penny


Cluster 3: Small- Dollars


Cluster 4: Medium-Hundreds


Cluster 5: Extra Large-Millions 